---
title: "linear model from Web"
author: "Mikaela Elias"
date: "Fall 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot

```{r}
plot(SIMS~ARM,data=data)
model.1 =lm(SIMS~ARM, data=data)
abline(model.1)
```
       
 The scatterplot shows that simulated strength score increases depending on the amount of arm strength exerted by an individual.     
          
### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
       
The numerical results shows that the correlation = 0.69 shows that there is a moderate positive correlation between simulated strength scores and arm strength. 
         
### Inferential  (Build model.1)

```{r}
summary.lm(model.1)
```
         
Model.1 shows that intercept -4.095 and coefficient 0.055 of ARM are both significantly different from 0. 
        
#### Predict at target point
      
```{r}
targetPoint = data.frame(GRIP=94, ARM=88)
predictInt = predict(model.1, targetPoint, interval = "prediction")
print(predictInt)
```
         
At ARM=88, GRIP=94, the predicted SIMS score is 0.71. 95% confidence interval for that score is SIMS between -1.73 and 3.14. 

#### scatterplot with model fit

           
## Model 2 SIM~GRIP

### Now add in scatterplot
      
```{r}
plot(SIMS~GRIP,data=data)
model.2 =lm(SIMS~GRIP, data=data)
abline(model.2)
```
         
The scatterplot shows that simulated strength score increases depending on the amount of GRIP strength exerted by an individual.      
       
### Numerical results 
      
```{r}
cor(SIMS~GRIP,data=data)
```
The numerical results shows that the correlation = 0.64 shows that there is a moderate positive correlation between simulated strength scores and arm strength.        
                  
### Inferential  (Build model.2)
         
```{r}
summary.lm(model.2)
```
         
Model.2 shows that intercept -4.809 and coefficient 0.045 of GRIP are both significantly different from 0.          
          
#### predict model.2 at target point
        
```{r}
targetPoint = data.frame(GRIP=94, ARM=88)
predictInt = predict(model.2, targetPoint, interval = "prediction")
print(predictInt)
```
         
At ARM=88, GRIP=94, the predicted SIMS score is -0.54. 95% confidence interval for that score is SIMS between -3.11 and 2.04.   


#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

```{r}
model.3 =lm(SIMS~ARM+GRIP, data=data)
summary (model.3)
```
          
Model.3 shows that intercept -5.434, coefficient 0.038 of ARM, and coefficient of GRIP 0.025 are all significantly different from 0.  

### Numerical results (cor)
         
```{r}
cor(data)
```
This shows the SIMS correlates with GRIP and ARM. 
         
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  
```{r}
predict(model.3, targetPoint, interval = "prediction")
```
At ARM=88, GRIP=94, the predicted SIMS score is 0.15. 95% confidence interval for that score is SIMS between -2.14 and 2.44.      
     
## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

```{r}
# Compare the two results models using ANOVA
anova_result <-anova(model.1, model.3)
print(anova_result)
```
       
The p-value of the comparison of model.1 and model.3 is 0.000004994 which is less than 0.5, this means that the models are different and model.3 is better.

```{r}
# Compare the two results models using ANOVA
anova_result <-anova(model.2, model.3)
print(anova_result)
```
    
The p-value of the comparison of model.2 and model.3 is 0.000000001495 which is less than 0.5, this means that the models are different and model.3 is better.